---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Download the data 
library(ggplot2)
library(dplyr)
#question1
```{r} 
if(!file.exists("./dati")){dir.create("./dati")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./dati/activity.zip",method="curl")
unzip(zipfile="./dati/activity.zip",exdir="./dati")
activity_data <- read.csv("./dati/activity.csv")
activity_data$date <- as.Date(activity_data$date)
```
## What is mean total number of steps taken per day?
#question2
```{r}
steps_days <- activity_data %>%
  group_by(date) %>%
  summarize(sumsteps = sum(steps, na.rm = TRUE))
head(steps_days,5)
hist(steps_days$sumsteps, xlab="Number of steps", main="Histogram")
mean <- round(mean(steps_days$sumsteps),digits = 2)
median <- round(median(steps_days$sumsteps),digits = 2)
print(paste("Mean:",mean))
print(paste("Median:",median))
```
## What is the average daily activity pattern?
#question3
```{r}
steps_interval <- activity %>%
  group_by(interval) %>%
  summarize(meansteps = mean(steps, na.rm = TRUE)) 
head(steps_interval,5)
plot(steps_interval$meansteps ~ steps_interval$interval, type="l", xlab = "5 Minute Intervals", ylab = "steps average")
print(paste("5-minute intrerval with maximum number of steps: ",steps_interval$interval[which.max(steps_interval$meansteps)]))
```
## Imputing missing values
#question4
```{r}
print(paste("Number of rows with NA: ",sum(is.na(activity$steps))))

activity_fill <- activity  
for (i in 1:nrow(activity)){
  if(is.na(activity$steps[i])){
    activity_fill$steps[i]<- steps_interval$meansteps[activity_fill$interval[i] == steps_interval$interval]
  }
}
```
steps_days <- activity_fill %>%
  group_by(date) %>%
  summarize(sumsteps = sum(steps, na.rm = TRUE)) 
hist(steps_days$sumsteps, xlab="number of steps")
mean_fill <- round(mean(steps_days$sumsteps), digits = 2)
median_fill <- round(median(steps_days$sumsteps), digits = 2)
print(paste("The mean is: ", mean(mean_fill)))
print(paste("The median is: ", mean(median_fill)))

##Are there differences in activity patterns between weekdays and weekends?
#question5
```{r}
DoW <- activity_fill
DoW$date <- as.Date(DoW$date)
DoW$day <- ifelse(weekdays(DoW$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
DoW$day <- as.factor(DoW$day)

WDay <- filter(DoW, DoW$day == "weekday")
Weekend <- filter(DoW, DoW$day == "weekend")

WDay <- WDay %>%
  group_by(interval) %>%
  summarize(steps = mean(steps)) 

WDay$day <- "weekday"

Weekend <- Weekend %>%
  group_by(interval) %>%
  summarize(steps = mean(steps)) 
Weekend$day <- "weekend"

wkdayWkend <- rbind(WDay, Weekend)
wkdayWkend$day <- as.factor(wkdayWkend$day)

r <- ggplot (wkdayWkend, aes (interval, steps))
r + geom_line() + facet_grid (day~.) + 
  theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14)) + 
  labs(y = "steps") + labs(x = "Interval") + 
  theme(plot.title = element_text(hjust = 0.5))
```

















